\section{Subgroups}

\subsection{Definitions and Basics}
Although we have defined subgroups in section $1$, we repeat the definition here.
\begin{definition}
Let $G$ be a group. A subset $H$ of $G$ is a subgroup of $G$ if $H$ is nonempty and it is closed under products and inverses. That is, $x,y\in H$ implies $x^{-1}\in H$ and $xy\in H$. If $H$ is a subgroup of $G$, we write $H\leq G$.
\end{definition}

If $H\leq G$ and $H\neq G$, we write $H<G$.

\begin{example}
    $\mathbb{Z}\leq\mathbb{Q}$ and $\mathbb{Q}\leq\mathbb{R}$ under the operation of addition.
    
    If $G=D_{2n}$, $H=\{1,r,r^2,\ldots,r^{n-1}\}$ is a subgroup of $G$.
\end{example}

Note that the relation $\leq$ is transitive. That is, if $K\leq H$ and $H\leq G$, then $K\leq G$.

\begin{theorem}[Subgroup Criterion]
A subset $H$ of a group $G$ is a subgroup if and only if
\begin{enumerate}
    \item $H\neq\emptyset$.
    \item for all $x,y\in H$, $xy^{-1}\in H$.
\end{enumerate}
Further, if $H$ is finite, then it suffices to check that $H$ is nonempty and is closed under multiplication.
\end{theorem}
\begin{proof}
    If $H\leq G$, the two given statements clearly hold as $H$ contains the identity of $G$ and is closed under inverses and multiplication.
    
    To prove the converse, let $x$ be any element of $H$ (which exists as $H\neq\emptyset$). We have $xx^{-1}\in H\implies 1\in H$. As $H$ contains $1$, for any element $h$ of $H$, $H$ contains $1h^{-1}=h^{-1}$, that is, it is closed under inverses. For any $x$ and $y$ in $H$, as $y^{-1}\in H$, we have that $x(y^{-1})^{-1}=xy\in H$, that is, $H$ is closed under multiplication.
    
    \vspace{1mm}
    To prove the second part, we see that $x,x^2,x^3,\ldots\in H$ for any $x\in H$. Using \ref{finGrpFinOrd}, we see that $x$ is of finite order $n$. Then $x^{-1}=x^{n-1}\in H$ so $H$ is closed under inverses.
\end{proof}

\begin{exercise}
    Let $G$ be a group and $H,K$ be subgroups of $G$. Show that $H\cup K$ is a subgroup if and only if $H\subseteq K$ or $K\subseteq H$.
\end{exercise}

\begin{exercise}
    Let $G$ be a group and $H,K$ be subgroups of $G$. Show that $H\cap K$ is also a subgroup of $G$.
\end{exercise}

\begin{exercise}
\label{subgrIntersubgr}
    Let $G$ be a group. Prove that the intersection of an arbitrary nonempty collection of subgroups of $G$ is again a subgroup of $G$.
\end{exercise}

\begin{exercise}
    Let $G$ be a group of order $n>2$. Show that $G$ cannot have a subgroup $H$ of order $n-1$.
\end{exercise}

\begin{exercise}
    Let $G$ be a group. Let $H=\{g\in G\mid |g|<\infty\}$. Show that $H\leq G$ if $G$ is abelian. In this case, $H$ is called the \textit{torsion subgroup} of $G$. Give an example where $G$ is non-abelian and $H$ is not a subgroup of $G$.
\end{exercise}

\begin{exercise}
    Let $H$ be a subgroup of $\mathbb{Q}$ under addition with the property that $\frac 1x\in H$ for every nonzero $x\in H$. Show that $H=\{0\}$ or $\mathbb{Q}$.
\end{exercise}

\subsection{Centralizers, Normalizers, Stabilizers and Kernels}

We now introduce some important subgroups.

\begin{definition}
    Let $G$ be a group and $A$ be any nonempty subset of $A$. Define
    $$C_G(A)=\{g\in G\mid gag^{-1}=a\text{ for all }a\in A\}.$$
    This subset is called the \textit{centralizer} of $A$ in $G$.
\end{definition}

Since $gag^{-1}=g$ if and only if $ga=ag$, $C_G(A)$ is the set of all elements that commute with every element of $A$.

Now observe that $C_G(A)$ is a subgroup of $G$ as first of all, $1\in C_G(A)$ so $C_G(A)\neq\emptyset$, and second of all, if $x,y\in C_G(A)$, we have $xax^{-1}=a$ and $yay^{-1}=a$, that is, $y^{-1}ay=a$ for all $a\in A$. We then have $a=xax^{-1}=x(y^{-1}ay)x^{-1}=(xy^{-1})a(xy^{-1})^{-1}$ so $xy^{-1}\in C_G(A)$. Thus, $C_G(A)\leq G$.

\begin{definition}
    Let $G$ be a group. Define
    $$Z(G)=\{g\in G\mid gx=xg\text{ for all }x\in G\}.$$
    This subset is called the \textit{center} of $G$.
\end{definition}

$Z(G)$ is the set of all elements that commute with every element of $G$.

As $Z(G)=C_G(G)$, we have $Z(G)\leq G$.

\begin{definition}
    Let $G$ be a group and $A$ be a subset of $G$. Define $gAg^{-1}=\{gag^{-1}\mid a\in A\}$. Define
    $$N_G(A)=\{g\in G\mid gAg^{-1}=A\}.$$
    This set is called the \textit{normalizer} of $A$ in $G$.
\end{definition}

The proof that $N_G(A)\leq G$ is similar to that we used to prove that $C_G(A)\leq G$.

Note that $C_G(A)\leq N_G(A)$.

\vspace{2mm}
If $G$ is an abelian group, $Z(G)=G$. Further, for any subset $A$ of $G$, $N_G(A)=C_G(A)=G$ as $gag^{-1}=gg^{-1}a=a$ for all $a\in A, g\in G$.

\begin{exercise}
    Show that the center of $D_8$ is $\{1,r^2\}$.
\end{exercise}

The fact that centralizers and normalizers are subgroups is in fact a special case of a results in group actions. 

We now introduce stablizers and kernels of group actions.

\begin{definition}
    Let $G$ be a group that acts on a set $S$. Let $s\in S$ be some fixed elements. Define
    $$G_s=\{g\in G\mid g\cdot s=s\}$$
\end{definition}

We shall now show that $G_s\leq G$. First of all, $1\in G_s$ by the definition of a group action. If $x,y\in G_s$, we have
\begin{align*}
    s &= 1\cdot s \\
      &= (x^{-1}x)\cdot s \\
      &= x^{-1}\cdot (x\cdot s) \\
      &= x^{-1}\cdot s
\end{align*}
so $x^{-1}\in G_s$ and
\begin{align*}
    (xy)\cdot s &= x\cdot(y\cdot s) \\
                &= x\cdot s \\
                &= s
\end{align*}
We see that $G_s$ is nonempty and is closed under inverses and multiplication. It is thus a subgroup of $G$.

Recall the definition of a \textit{kernel} of an action, \ref{defKerGrpAc}. Using \ref{subgrIntersubgr} and the fact that $G_s\leq G$ for all $s\in S$ yields the result that the kernel of any group action is a subgroup of the group.

\vspace{3mm}
We now see that $C_G(A)$ is merely the kernel of the group action of $G$ acting on $A$ as $g\cdot a= gag^{-1}$ (so it is a subgroup of $G$) and $N_G(A)$ is the stabilizer of the group action of $G$ acting on $\mathcal{P}(A)$ (the power set of $A$) as $g\cdot A=gAg^{-1}$ (so it is a subgroup of $G$).

\begin{exercise}
    Prove that $C_G(Z(G))=N_G(Z(G))=G$.
\end{exercise}

\begin{exercise}
    Prove that $H\leq N_G(H)$ for a subgroup $H$ of a group $G$.
\end{exercise}

\begin{exercise}
    For any subgroup $H$ of group $G$ and subset $A$ of $G$, define $N_H(A)=\{h\in H\mid hAh^{-1}=A\}$. Prove that $N_H(A)=N_G(A)\cap H$ and deduce that $N_H(A)\leq N_G(A)$.
\end{exercise}

\begin{exercise}
    Let $F$ be a field and the Heisenberg group $H(F)$ be defined as in \ref{defOfHeisenbergGrp}. Determine $Z(H(F))$ and prove that $Z(H(F))\cong (F,+)$.
\end{exercise}

\subsection{Cyclic Groups and Cyclic Subgroups}

\begin{definition}
    A group $H$ is \textit{cyclic} if there is some element $x\in H$ such that $H=\{x^n\mid n\in\mathbb{Z}\}$.
\end{definition}

In this case we write $H=\langle x\rangle$ and say that $H$ is \textit{generated} by $x$ and $x$ is a generator of $H$. The generator of a cyclic group need not be unique (as if $x$ is a generator, so is $-x$).

Note that any cyclic group is abelian.

\begin{example}
    The group $(\mathbb{Z},+)$ is generated by $1$ (here $1$ is the integer $1$ and not the identity).
\end{example}

\begin{theorem}
    Let $H=\langle x\rangle$. Then $|H|=|x|$ (where if one side of the inequality is infinite, so is the other).
\end{theorem}

\begin{proof}
    This proof is trivial and is left as an exercise to the reader.
\end{proof}

It is observed that there is a great deal of similarity between $H=\langle x\rangle$, where $|x|=n$, and $\mathbb{Z}/n\mathbb{Z}$. Both of them appear to have very similar structure. It turns out that these two groups are isomorphic, which we shall prove shortly. First, let us prove the following.

\begin{theorem}
    Let $G$ be an arbitrary group, $x\in G$, and let $m,n\in\mathbb{Z}$. If $x^n=1$ and $x^m=1$, then $x^d=1$, where $d=(m,n)$. In particular, if $x^m=1$ for some $m\in\mathbb{Z}$, then $|x|\mid m$.
\end{theorem}

\begin{proof}
    By the Euclidean algorithm, there exist integers $r$ and $s$ such that $d=mr+ns$. We have
    $$x^d=x^{mr+ns}=(x^m)^r(x^n)^s=1.$$
    This proves our first claim.
    
    Next, let $n=|x|$ and $x^m=1$. We have $x^d$=1, where $d=(|x|,m)$. Note that $0<d\leq |x|$ and $|x|$ is the smallest positive integer $k$ such that $x^k=1$. This implies that $d=|x|$ and $|x|=(|x|,m)$. Thus, $|x|\mid m$. 
\end{proof}

\begin{theorem}
    Any two cyclic groups of the same order are isomorphic. More specifically,
    \begin{enumerate}
        \item if $n\in\mathbb{Z}^+$ and $H=\langle x\rangle$ and $K=\langle y\rangle$ are both of order $n$, $H\cong K$.
        \item if $\langle x\rangle$ is an infinite cyclic group, $(\mathbb{Z},+)\cong\langle x\rangle$.
    \end{enumerate}
\end{theorem}

\clearpage